---
title: "austin crime viz"
author: "@nntrn"
date: "10/25/2018"
output:
    html_notebook:
      highlight: textmate
      code_folding: none
      css: "/Users/annie/programming-notes/resources/Rnotebook-style.css"
      favicon: "/Users/annie/programming-notes/resources/favicon.png"
      df_print: kable
      mathjax: null
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

knitr::opts_knit$set(root.dir = "/Users/annie/Desktop/notes/R/austin crime/")
knitr::opts_chunk$set(fig.width = 6, fig.height = 3)

library(lubridate)
library(kableExtra)
library(DescTools)
library(tidyverse)

my_theme <- theme_minimal() +
  theme(
    axis.text.x = element_text(colour = "grey20", size = 13, face = "bold"),
    axis.text.y = element_text(colour = "grey20", size = 8, face = "bold"),
   legend.position="bottom", legend.box = "horizontal"
  )

theme_set(my_theme)

```

austin crime data for 2017

> crime is at the lowest in january (this is also true for 2018)

# setting up


get the data here:

* [2017](https://data.austintexas.gov/Public-Safety/Crime-Reports-2017/4bxg-n3iv)
* [2018](https://data.austintexas.gov/Public-Safety/Crime-Reports-2018/vmn9-3bvu)


```{r cdata, results = 'hide'}
crime <- read_csv("Crime_Reports_2017.csv",
  col_types = cols(
    `Occurred Date` = col_date("%m/%d/%Y"),
    `Report Date` = col_date("%m/%d/%Y")
  )
)
```

```{r save, results = 'hide', echo=FALSE, include=FALSE}
write.csv(crime2, file = "/Users/annie/programming-notes/R/austin-crime/output/crime2.csv",row.names = FALSE)    
write.csv(reported_streets, file = "/Users/annie/programming-notes/R/austin-crime/output/reported_streets.csv",row.names = FALSE)  

write.csv(p, file = "/Users/annie/programming-notes/R/austin-crime/output/p.csv",row.names = FALSE)  
```


# data manipulation

keep this handy:
`round_to <- function(x, to = 10) round(x / to) * to`

```{r manipulate, results='hide'}
# make column names R friendly
colnames(crime) <- gsub(" ", "_", colnames(crime)) %>% tolower()

## ROUND TO FUNCTION 
round_to <- function(x, to = 10) round(x / to) * to

crime2 <- crime %>%
    filter(occurred_date >= as.Date("2017-01-01")) %>%
    mutate(
    ## TIME RELATED ======
    month = ordered(toupper(format(occurred_date, "%b")), toupper(month.abb)),
    day = ordered(
    toupper(format(occurred_date, "%a")),
    c("SUN", "MON", "TUE", "WED", "THU", "FRI", "SAT")
    ),
    hoc_parent = round_to(highest_offense_code, 100),
    hour = round_to(occurred_time, 100),
    
    ## LOCATION RELATED ========
    street = sapply(address, function(x) {
    gsub(pattern = " BLOCK ", replacement = " ", x)
    }),
    
    street_num = sapply(street, function(x) {
    str_extract(x, "^[0-9]*")
    }),
    
    # removes all leading street numbers
    street_name = sapply(street, function(x) {
    gsub("^[0-9]*\\s", "", x)
    }),
    
    street_name1 = sapply(street_name, function(x) {
    str_split_fixed(x, " / ", 2)[1]
    }),
    
    street_name2 = sapply(street_name, function(x) {
    str_split_fixed(x, " / ", 2)[2]
    }),
    
    ## value is split for incidents with intersections
    value = 1,
    value = ifelse(street_name2 != "", .5, value)
    )
    
reported_streets <-
  rbind(
    crime2 %>%
      select(id = incident_number, street_name = street_name2, zip = zip_code, value) %>%
      filter(street_name != ""),
    crime2 %>%
      select(id = incident_number, street_name = street_name1, zip = zip_code, value) %>%
      filter(street_name != "")
    )

## top crime reported streets
reported_streets %>%
  group_by(street_name, zip) %>%
  summarise(n = n(), s = sum(value)) %>%
  arrange(desc(n))

```

## streets to watch out for

```{r saved}

p <- reported_streets %>%
  left_join(
      crime2 %>%
          select(id = incident_number, category_description, month, zip_code, occurred_date, highest_offense_description), 
      by = "id"
      ) 

p <- p %>%
    filter(street_name != "UNKNOWN") %>%
    group_by(highest_offense_description, street_name) %>%
    summarise(n = n(), s = sum(value)) %>%
    arrange(desc(n))

# get top 25 most committed offenses
top25 <- p %>%
    filter(
        street_name != "UNKNOWN"
        ) %>%
    group_by(highest_offense_description) %>%
    summarise(n = n()) %>%
    arrange(desc(n)) %>% 
    top_n(25)


## rank offense by group
rank <- p %>%
    group_by(highest_offense_description) %>%
    mutate(my_ranks = order(order(n, decreasing=TRUE))) %>%
## filter to get top 10 streets of the top 25 most committed offenses 
    filter(
        my_ranks <= 10,
        highest_offense_description %in% top25$highest_offense_description
        ) %>% 
    arrange(
        highest_offense_description,
        my_ranks
    ) %>%
    select(rank=my_ranks, highest_offense_description, street_name, n, s)


input_street <- "BURNET RD"

rank %>%
ggplot(aes(highest_offense_description,as.ordered(rank),label=street_name)) + geom_point(color="white") +
     geom_text(hjust=0,vjust=0,size=2) +
  theme_void()+
  theme(
    axis.text.x = element_text(colour = "grey20", size = 8, face = "bold"),
    axis.text.y = element_text(colour = "grey20", size = 8, face = "bold"),
   legend.position="bottom", legend.box = "horizontal"
  ) + coord_flip()

offense_list <- c("ASSAULT BY THREAT", "DWI", "BURGLARY OF RESIDENCE","CRIMINAL MISCHIEF", "HARASSMENT", "POSS MARIJUANA")

rank %>%
    filter(
        highest_offense_description %in% offense_list
        ) %>%
    ggplot(aes(n, rev(as.ordered(rank)), label=street_name)) + 
    geom_point(size=3, aes(color=n,alpha=.7)) +
    facet_grid(.~highest_offense_description) +
    geom_text(x=25,hjust=0,vjust=0,size=3) + 
    theme_light() + theme(axis.text.y = element_blank()) +
    geom_text(aes(label=ifelse(street_name==input_street,input_street,'')),x=25,hjust=0,vjust=0,size=3,color="red") 


```

### how to rank by group 

```r
df %>%
    group_by(customer_name) %>%
    mutate(my_ranks = order(order(order_values, order_dates, decreasing=TRUE)))
```

[source](https://stackoverflow.com/questions/31859175/how-to-rank-within-groups-in-r)

**note:**  `n` is the number of occurence a street appeared (including intersection), `s` is the "value" of those occurences

## neighborhood data

```{r ggplot}

input_zip <- "78757"
input_street <- "E RIVERSIDE DR"

p <- reported_streets %>%
  left_join(
      crime2 %>%
          select(id = incident_number, category_description, month, zip_code, occurred_date), 
      by = "id"
      ) %>%
  filter(
    #street_name == input_street,
    zip_code == input_zip,
    category_description != ""
  )

# plot for crime by month for specified zipcode
p %>%
  group_by(month, category_description) %>%
  summarise(n = n(), s = sum(value)) %>%
  ggplot(aes(month, s, color = category_description)) +
  geom_point(size=2,alpha =.7) + 
    # labs(title = paste("2017 crime for all of ATX"))
     labs(title = paste("2017 crime for ",input_street))

# plot to get street count 
p %>%  group_by(street_name, category_description) %>%
  summarise(n = n()) %>% 
    filter(n > 3) %>% #top_n(15) %>%
    ggplot(aes(x=n, y=street_name, colour=category_description, label=n)) +
    geom_point(shape=16,size=2) + 
    geom_text(aes(label=ifelse(n>6,n,'')),hjust=-1,vjust=.25,size=3) +
    labs(title = paste("crime in ",input_zip))



```


## polar coordinates

```{r polar}
mtcars$car = row.names(mtcars)

p %>%
ggplot(aes(x=month, y=value, fill=highest_offense_description)) +
    geom_bar(stat="identity") + theme_light() +
#    scale_fill_gradient(low=’red’, high=’white’, limits=c(5,40)) +
    theme(axis.title.y=element_text(angle=0)) +
    theme(axis.text.x = element_text(angle=45, vjust = 1, hjust=1)) + coord_polar()

```
